Pyg4ometry: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-08-29 public/paper/pyg4ometry.tex
Updated paper diagrams.

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage[T1]{fontenc}
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{siunitx}
\usepackage{lineno}
\modulolinenumbers[5]
\usepackage{xcolor}
\usepackage{cleveref}

@@

\linenumbers

@@

\section{Geometric descriptions} \label{sec:geometric}
Central to description of a geometric description of a system is how a solid is defined in three dimensions. There are numerous different ways to describe a
geometry and these are described briefly in the following sections.

@@

{\small
\begin{verbatim}
import pyg4ometry.gdml as gd
import pyg4ometry.geant4 as g4
import pyg4ometry.visualisation as vi

# create empty data storage structure
reg = g4.Registry()

# expressions 
wx = gd.Constant("wx","100",reg)
wy = gd.Constant("wy","100",reg)
wz = gd.Constant("wz","100",reg)

@@

# materials
bm = g4.Material("G4_Galactic") 
wm = g4.Material("G4_Iron")

# solids
wb = g4.solid.Box("wb",wx,wy,wz,reg)
b  = g4.solid.Box("b",bx,by,bz)

# structure 
wl = g4.LogicalVolume(wb, wm, "wl", reg)
bl = g4.LogicalVolume(b, bm, "b", reg)
bp1 = g4.PhysicalVolume([0,0,0],
                        [0,0,0], 
                        bl, "b_pv1", wl) 
bp2 = g4.PhysicalVolume([0,0,0],
                        [-2*bx,0,0], 
                        bl, "b_pv2", wl)  
bp3 = g4.PhysicalVolume([0,0,0],
                        [2*bx,0,0], 
                        bl, "b_pv3", wl)

# gdml output
w = gd.Writer()
w.write(reg,"output.gdml")

# visualisation 
v = vi.VtkViewer()
v.addLogicalVolume(wl)
v.view()
\end{verbatim}
}

@@

The FLUKA CSG ASCII format is parsed using an ANTLR4-generated parser,
producing an abstract syntax tree (AST).  The resulting AST is then
walked to populate Region instances with zones and bodies.  An example
FLUKA region with the corresponding abstract syntax tree is shown in
Figure~\ref{fig:region-ast}.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=0.95\textwidth]{./model-scene/faradayCup2.pdf}
\caption{Example conversion of a simple FLUKA geometry to GDML, a) 
the original FLUKA geometry displayed in FLAIR, b) the GDML geometry 
viewed in pyg4ometry. The example is a Faraday cup used to capture 
and measure accelerator beam charge}
\label{fig:region-ast}
\end{center}
\end{figure*}

@@

\section{CAD to GDML conversion }
STEP and IGES files can be loaded into pyg4ometry, via an interface based on FreeCAD. FreeCAD is an open source CAD/CAM program, which in turn 
is based on OpenCASCADE. A STEP CAD file could be considered as a hierarchical tree of {\it part assemblies} and {\it part features}, where a part assembly is a 
collection of part features. A part feature can be used to create  a triangular mesh which can used to create a pyg4ometry tessellated solid. Assignment of materials 
and visualisation attributes must be performed by the user after conversion to GDML as rarely will CAD/CAM packages included the detailed information required 
for RT codes.

@@

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage[T1]{fontenc}
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{siunitx}
\usepackage{lineno}
\modulolinenumbers[5]
\usepackage{xcolor}
\usepackage{cleveref}
\usepackage{listings}

\lstset{
  showstringspaces=false,
  basicstyle=\ttfamily,
  columns=fullflexible,
%  frame=single,
  breaklines=true,
  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}

@@

% \linenumbers

@@

\section{Computer descriptions of geometry} \label{sec:geometric}
Central to a geometric description of a system is how a solid is defined in three dimensions. There are numerous different ways to describe a
geometry and these are described briefly in the following sections.

@@

\begin{lstlisting}[caption={A simple python script using pyg4ometry to create a simple Geant4 geometry.},label={lst:pythonRapidModelling}, language=Python]
# Import modules 
import pyg4ometry.gdml as gd
import pyg4ometry.geant4 as g4
import pyg4ometry.visualisation as vi

# Create empty data storage structure
reg = g4.Registry()

# Expressions 
wx = gd.Constant("wx","100",reg)
wy = gd.Constant("wy","100",reg)
wz = gd.Constant("wz","100",reg)

@@

# Materials
bm = g4.Material("G4_Galactic") 
wm = g4.Material("G4_Iron")

# Solids
wb = g4.solid.Box("wb",wx,wy,wz,reg)
b  = g4.solid.Box("b",bx,by,bz)

# Structure 
wl = g4.LogicalVolume(wb, wm, "wl", reg)
bl = g4.LogicalVolume(b, bm, "b", reg)
bp1 = g4.PhysicalVolume([0,0,0],[0,0,0],bl,"b_pv1", wl) 
bp2 = g4.PhysicalVolume([0,0,0],[-2*bx,0,0],bl,"b_pv2", wl)  
bp3 = g4.PhysicalVolume([0,0,0],[2*bx,0,0],bl,"b_pv3", wl)

# GDML output
w = gd.Writer()
w.write(reg,"output.gdml")

# Visualisation 
v = vi.VtkViewer()
v.addLogicalVolume(wl)
v.view()
\end{lstlisting}

@@

The FLUKA CSG ASCII format is parsed using an ANTLR4-generated parser,
producing an abstract syntax tree (AST).  The resulting AST is then
walked to populate Region instances with zones and bodies.  An example
FLUKA region with the corresponding abstract syntax tree is shown in
Figure~\ref{fig:region-ast}.
\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/faradayCup2.pdf}
\caption{Example conversion of a simple FLUKA geometry to GDML, a) 
the original FLUKA geometry displayed in FLAIR, b) the GDML geometry 
viewed in pyg4ometry. The example is a Faraday cup used to capture 
and measure accelerator beam charge}
\label{fig:region-ast}
\end{center}
\end{figure*}

@@

\section{CAD/STL to GDML conversion }
STEP and IGES files can be loaded into pyg4ometry, via an interface based on FreeCAD. FreeCAD is an open source CAD/CAM program, which in turn 
is based on OpenCASCADE. A STEP CAD file could be considered as a hierarchical tree of {\it part assemblies} and {\it part features}, where a part assembly is a 
collection of part features. A part feature can be used to create  a triangular mesh which can used to create a pyg4ometry tessellated solid. Assignment of materials 
and visualisation attributes must be performed by the user after conversion to GDML as rarely will CAD/CAM packages included the detailed information required 
for RT codes.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/SectorBend.pdf}
\caption{CAD-GDML}
\label{fig:cad-gdml}
\end{center}
\end{figure*}

@@

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/model.pdf}
\caption{CAD-GDML}
\label{fig:model}
\end{center}
\end{figure*}
